---
title: "Northeast U.S."
output: html_notebook
---
```{r setup, include = FALSE, message=FALSE}
# load libraries
library(EML)
library(emld)
library(tidyverse) # for data table wrangling
library(knitr) # for web presentation of tables
library(kableExtra) # for web presentation of tables

# load data

# creates an object called spp
load(here::here("data_raw", "neus_SVSPP.RData"))

# creates an object called survdat
load(here::here("data_raw", "neus_Survdat.RData"))

strata <- read_csv("https://raw.githubusercontent.com/pinskylab/OceanAdapt/master/data_raw/neus_strata.csv", n_max = 6)

column_defs <- readRDS(here::here("metadata","column-definitions.rds"))
```

## Source:
[NOAA Northeast Fisheries Science Center Spring and Fall Bottom Trawl Surveys](http://www.nefsc.noaa.gov/femad/ecosurvey/mainpage/)

## Related information:  
- Source data can be found here: [Fall](https://inport.nmfs.noaa.gov/inport/item/22560) [Spring](https://inport.nmfs.noaa.gov/inport/item/22561).  
- [ Guide to some trawl-caught marine fishes from Maine to Cape Hatteras, North Carolina, Flescher, 1980](https://spo.nmfs.noaa.gov/content/circular-431-guide-some-trawl-caught-marine-fishes-maine-cape-hatteras-north-carolina)  
- [A brief historical review of the Woods Hole Laboratory trawl survey time series, Azarovitz, 1981](http://dmoserv3.whoi.edu/data_docs/NEFSC_Bottom_Trawl/Azarovitz1981.pdf)  
- [An Evaluation of the Bottom Trawl Survey Program of the Northeast Fisheries Center, NEFSC, 1988](https://www.st.nmfs.noaa.gov/tm/nec_image/nec052image.pdf)  
- [A historical perspective on the abundance and biomass of northeast demersal complex stocks from NMFS and Massachusetts inshore bottom trawl surveys, 1963-2002, Sosebe and Cardin, 2006](https://repository.library.noaa.gov/view/noaa/5259)   
- [Estimation of Albatross IV to Henry B. Bigelow Calibration Factors, Miller et al., 2010](https://www.nefsc.noaa.gov/publications/crd/crd1005/crd1005.pdf)
- [Northeast Fisheries Science Center Bottom Trawl Survey Protocols for the NOAA Ship Henry B. Bigelow, Politis et al., 2014](https://www.nefsc.noaa.gov/publications/crd/crd1406/)  
- [Groundfish Bottom Trawl Survey Protocols, NOAA Fisheries, 2018](https://www.fisheries.noaa.gov/resource/document/groundfish-bottom-trawl-survey-protocols)  
- [Density-Independent and Density-Dependent Factors Affecting Spatio-Temporal Dynamics of Atlantic Cod (Gadus Morhua) Distribution in the Gulf of Maine, Zengguang et al., 2018](https://doi.org/10.1093/icesjms/fsx246)
- [Technical Documentation, State of the Ecosystem Report, NEFSC, 2019](https://noaa-edab.github.io/tech-doc/)  
- [NEFSC trawl strata](neus_NEFSC_trawl_strata.pdf)  

## How we process the data:  
- Before 2020, we emailed a staff member at NOAA with a data request and recieved a RData file.  This file was a combination of the SVBIO, SVCAT, and SVSTA files and some column names were changed.  Now we download the files from the publicly available data set.  We combine those files and change the column names to match the column names we used to receive so that subsequent code will work.  The changes include changing EST_YEAR to YEAR, changing DECDEG_BEGLAT to LAT, DECDEG_BEGLON to LON, AVGDEPTH to DEPTH, EXPCATCHWT to BIOMASS.  
- There are some commas and special characters in the svcat.csv files that cause them to parse incorrectly.  We import those files with read_lines, remove the commas and special characters from the comments, and proceed to read them into R as .csvs.  
- We group the data by YEAR, SEASON, LAT, LON, DEPTH, CRUISE6, STATION, STRATUM, and SVSPP and sum the BIOMASS (which is reported by sex) to calculate wtcpue.  
- We create a haulid by combining a 6 digit leading zero cruise number with a 3 digit leading zero station number and a 4 digit leading zero stratum number, separated by "-", for example: (cruise-station-stratum) 456354-067-0001.  
- We convert square nautical miles to square kilometers.  
- We remove any SCINAME spp values that contain the word "egg" or "unidentified", or where the only value in the SCINAME field is white space.  
- We group the data by haulid, stratum, stratumarea, year, lat, lon, depth,  and spp and then sum up all of the wtcpue values for each group and reassign that as the wtcpue.
- We separate the trawls into Fall and Spring seasons.

## What the raw data includes:  
The current files of raw data for the Northeast U.S. are `r list.files(here::here("data_raw"), pattern = "neus")`.  neus_svspp.csv contains the same data as neus_SVSPP.RData.

### neus_strata.csv is constant through the years.
- In 2020 this file was updated to add a leading zero to the STRATUM column to match the STRATUM column in the publicly available data.  
```{r include = FALSE}
st_attributes <- tibble(
  attributeName = dput(names(strata))) %>% 
  left_join(column_defs)
```
```{r echo=FALSE, message=TRUE, warning=TRUE}
knitr::kable(st_attributes) %>% 
  kableExtra::kable_styling(latex_options = "scale_down")
```

### neus_spring_svcat.csv and neus_fall_svcat.csv are updated annually.
```{r include = FALSE}
# create a table with each column from the dataset represented as a row.  There will be an error message about undefined units but they are defined later when the dataset is created.
attributes <- tibble(
  attributeName = dput(names(neus_spr_catch))) %>% 
  left_join(column_defs) %>% 
  mutate(formatString = NA,
         formatString = ifelse(attributeName == "YEAR", unit, formatString), 
         formatString = ifelse(attributeName == "DATETIME", unit, formatString)) %>% 
  rename(col_classes = col_class)




dput(attributes$col_classes)

# attributeList <- set_attributes(attributes)
```
```{r echo=FALSE}
knitr::kable(select(attributes, -formatString)) %>% 
  kableExtra::kable_styling(latex_options = "scale_down")
```
### neus_spring_svsta.csv and neus_fall_svsta.csv are updated annually.
```{r include = FALSE}
# create a table with each column from the dataset represented as a row.  There will be an error message about undefined units but they are defined later when the dataset is created.
attributes <- tibble(
  attributeName = dput(names(neus_spr_station))) %>% 
  left_join(column_defs) %>% 
  mutate(formatString = NA,
         formatString = ifelse(attributeName == "YEAR", unit, formatString), 
         formatString = ifelse(attributeName == "DATETIME", unit, formatString)) %>% 
  rename(col_classes = col_class)

dput(attributes$col_classes)

# attributeList <- set_attributes(attributes)
```
```{r echo=FALSE}
knitr::kable(select(attributes, -formatString)) %>% 
  kableExtra::kable_styling(latex_options = "scale_down")
```




## neus_svspp.csv connects species codes to species names.
```{r include = FALSE}
spp_attributes <- tibble(
  attributeName = dput(names(spp))) %>% 
  left_join(column_defs) %>% 
  mutate(formatString = NA,
         formatString = ifelse(attributeName == "YEAR", unit, formatString), 
         formatString = ifelse(attributeName == "DATETIME", unit, formatString))

dput(spp_attributes$col_class)

# attributeList <- set_attributes(spp_attributes, col_classes = c("character", "character", "character", "character", "character"))
```
```{r echo=FALSE}
knitr::kable(select(spp_attributes, -formatString)) %>% 
  kableExtra::kable_styling(latex_options = "scale_down")
```


```{r include = FALSE}
# define custom units
id <- c("kilograms per tow", "number per tow", "square kilometers")
dimension <- c("weight", "count", "area")
unitTypes <- data.frame(
  id = id, dimension = dimension, stringsAsFactors = FALSE
)
units <- data.frame(
  id = id, unitType = unitTypes, stringsAsFactors = FALSE
)

unitList <- set_unitList(units, unitTypes)

# define the rest of the data
# <title>
	# character vector title
	dataTitle <- "Northeast U.S. bottom trawl survey"

# <creator>
	# organization name
	afsc.name <- "National Oceanic and Atmospheric Administration (NOAA) Alaska Fisheries Science Center (AFSC) Resource Assessment and Conservation Engineering Division (RACE)"
	# create organization address
	afsc_address <- list( 
		deliveryPoint = "7600 Sand Point Way, N.E. bldg. 4",
		city = "Seattle",
		administrativeArea = "WA",
		postalCode = "98115",
		country = "USA"
	)
	
	publisher <- list(
	organizationName = afsc.name,
	address = afsc_address,
	onlineUrl = "http://www.afsc.noaa.gov/RACE/groundfish/survey_data/default.htm")

	R_person <- person(given = "Wayne", family = "Palsson", email = "Wayne.Palsson@noaa.gov", role = "cre")
	wayne <- as_emld(R_person)

	metadataProvider <- as_emld(person(given="Michelle", family = "Stuart",email = "michelle.stuart@rutgers.edu"))
	
others <- as.person("Bob Lauth [ctb] <Bob.Lauth@noaa.gov>")
associatedParty <- as_emld(others)
associatedParty[[1]]$organizationName <- afsc.name
	
# <pubDate>
pubDate <- "2012"

# <intellectualRights>
	# IR <- "this is just a long piece of text"

# <abstract>
abstract <- set_TextType(text = "The Resource Assessment and Conservation Engineering Division (RACE) of the Alaska Fisheries Science Center (AFSC) conducts bottom trawl surveys to monitor the condition of the demersal fish and crab stocks of Alaska. These data include catch per unit effort for each identified species at a standard set of stations. This is a subset of the main racebase datase. Excluded are non standard stations, earlier years using different gear, and other types of data collected other than species id, species weight, water temperature and depth.")

# This isn't working right now
# # <methods>
# 	methods <- 	list(dataSource="http://www.afsc.noaa.gov/RACE/groundfish/survey_data/default.htm",
# 				description="http://www.afsc.noaa.gov/RACE/groundfish/survey_data/metadata_template.php?fname=RACEweb.xml"
# 			) 
# 	set_methods(methods)

	

# ======================
# = Create EML Objects =
# ======================

neus.data <- set_physical("https://raw.githubusercontent.com/pinskylab/OceanAdapt/master/data_raw/ai1983_2000.csv")
# Create "dataTable": the physical data set with associated column definitions and units
neus.dataTable <- list(
		physical=neus.data,
		entityName="neus",
		entityDescription = "Bottom Trawl Data in the Northeast U.S.",
		attributeList = attributeList
)


eml <- list(
           packageId = uuid::UUIDgenerate(),  
           system = "uuid",
           dataset = list(
               title = dataTitle,
               creator = wayne,
               pubDate = pubDate,
               # abstract = abstract,
               contact = wayne,
               additionalMetadata = list(metadata = list(unitList = unitList)),
               dataTable = neus.dataTable
               ))


# validate
eml_validate(eml)

# Write EML data file
write_eml(eml, file=here::here("metadata","neus.xml"))
```

<!-- The Ecological Metadata Language file can be accessed [here](https://github.com/pinskylab/OceanAdapt/blob/master/metaData/neus.xml) -->





