---
title: "Gulf of Alaska"
output: html_notebook
---
```{r setup, include = FALSE, message=FALSE}
# load libraries
library(EML)
library(emld)
library(tidyverse) # for data table wrangling
library(knitr) # for web presentation of tables
library(kableExtra) # for web presentation of tables

# load data
sample_data <- read_csv("https://raw.githubusercontent.com/pinskylab/OceanAdapt/master/data_raw/goa1984_1987.csv", n_max = 6)

strata <- read_csv("https://raw.githubusercontent.com/pinskylab/OceanAdapt/master/data_raw/goa_strata.csv", n_max = 6)

column_defs <- readRDS(here::here("metadata","column-definitions.rds"))
```
## Source:
[NOAA Alaska Fisheries Science Center Groundfish Assessment Program surveys](http://www.nwfsc.noaa.gov/research/divisions/fram/groundfish/bottom_trawl.cfm)
[Download](https://www.afsc.noaa.gov/RACE/groundfish/survey_data/)

## Related papers:  

- [Data Report: 2009 Gulf of Alaska Bottom Trawl Survey](https://www.afsc.noaa.gov/Publications/AFSC-TM/NOAA-TM-AFSC-208.pdf)  
- [Groundfish Bottom Trawl Survey Protocols](https://www.fisheries.noaa.gov/resource/document/groundfish-bottom-trawl-survey-protocols)

## How we process the data:  
- Some of the files contain extra headers in the data rows, so we remove any data rows that contain the word "LATITUDE" in the LATITUDE column.  
- We create a haulid by combining a 3 digit leading zero vessel number with a 3 digit leading zero cruise number and a 3 digit leading zero haul number, separated by "-", for example: (vessel-cruise-haul) 354-067-001.  
- If wtcpue is recorded as "-9999", we change the value to NA.  
- We remove any SCIENTIFIC spp values that contain the word "egg" or where the only value in the SCIENTIFIC field is white space.  
- Any values SCIENTIFIC values that contain the word "Lepidopsetta" are changed to "Lepidopsetta sp." because more than one genus/spp combo was used to describe the same organism over time.  This also holds true for Myoxocephalus sp. excluding scorpius and Bathyraja sp. excluding panthera.
- We group the data by haulid, stratum, stratumarea, year, lat, lon, depth,  and spp and then sum up all of the wtcpue values for each group and reassign that as the wtcpue.  

## What the raw data includes:  
The current files of raw data for the Gulf of Alaska are `r list.files(here::here("data_raw"), pattern = "goa")`.

### goa_strata.csv is constant through the years with the column definitions  
```{r include = FALSE}
st_attributes <- tibble(
  attributeName = dput(names(strata))) %>% 
  left_join(column_defs)
```
```{r echo=FALSE, message=TRUE, warning=TRUE}
knitr::kable(st_attributes) %>% 
  kableExtra::kable_styling(latex_options = "scale_down")
```

### The remaining files are trawl data files, data added annually, with the column definitions
```{r include = FALSE}
# create a table with each column from the dataset represented as a row.  There will be an error message about undefined units but they are defined later when the dataset is created.
attributes <- tibble(
  attributeName = dput(names(sample_data))) %>% 
  left_join(column_defs) %>% 
  mutate(formatString = NA,
         formatString = ifelse(attributeName == "YEAR", unit, formatString), 
         formatString = ifelse(attributeName == "DATETIME", unit, formatString))

attributeList <- set_attributes(attributes, col_classes = c("numeric", "numeric", "character", "numeric", "Date", "Date", "numeric", "numeric", "character", "character", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric"))
```
```{r echo=FALSE}
knitr::kable(select(attributes, -formatString)) %>% 
  kableExtra::kable_styling(latex_options = "scale_down")
```

```{r include = FALSE}
# define custom units
id <- c("kilograms per tow", "number per tow", "square kilometers")
dimension <- c("weight", "count", "area")
unitTypes <- data.frame(
  id = id, dimension = dimension, stringsAsFactors = FALSE
)
units <- data.frame(
  id = id, unitType = unitTypes, stringsAsFactors = FALSE
)

unitList <- set_unitList(units, unitTypes)

# define the rest of the data
# <title>
	# character vector title
	dataTitle <- "Gulf of Alaska bottom trawl survey"

# <creator>
	# organization name
	afsc.name <- "National Oceanic and Atmospheric Administration (NOAA) Alaska Fisheries Science Center (AFSC) Resource Assessment and Conservation Engineering Division (RACE)"
	# create organization address
	afsc_address <- list( 
		deliveryPoint = "7600 Sand Point Way, N.E. bldg. 4",
		city = "Seattle",
		administrativeArea = "WA",
		postalCode = "98115",
		country = "USA"
	)
	
	publisher <- list(
	organizationName = afsc.name,
	address = afsc_address,
	onlineUrl = "http://www.afsc.noaa.gov/RACE/groundfish/survey_data/default.htm")

	R_person <- person(given = "Wayne", family = "Palsson", email = "Wayne.Palsson@noaa.gov", role = "cre")
	wayne <- as_emld(R_person)

	metadataProvider <- as_emld(person(given="Michelle", family = "Stuart",email = "michelle.stuart@rutgers.edu"))
	
others <- as.person("Bob Lauth [ctb] <Bob.Lauth@noaa.gov>")
associatedParty <- as_emld(others)
associatedParty[[1]]$organizationName <- afsc.name
	
# <pubDate>
pubDate <- "2012"

# <intellectualRights>
	# IR <- "this is just a long piece of text"

# <abstract>
abstract <- set_TextType(text = "The Resource Assessment and Conservation Engineering Division (RACE) of the Alaska Fisheries Science Center (AFSC) conducts bottom trawl surveys to monitor the condition of the demersal fish and crab stocks of Alaska. These data include catch per unit effort for each identified species at a standard set of stations. This is a subset of the main racebase datase. Excluded are non standard stations, earlier years using different gear, and other types of data collected other than species id, species weight, water temperature and depth.")

# This isn't working right now
# # <methods>
# 	methods <- 	list(dataSource="http://www.afsc.noaa.gov/RACE/groundfish/survey_data/default.htm",
# 				description="http://www.afsc.noaa.gov/RACE/groundfish/survey_data/metadata_template.php?fname=RACEweb.xml"
# 			) 
# 	set_methods(methods)

	

# ======================
# = Create EML Objects =
# ======================

goa.data <- set_physical("https://raw.githubusercontent.com/pinskylab/OceanAdapt/master/data_raw/ai1983_2000.csv")
# Create "dataTable": the physical data set with associated column definitions and units
goa.dataTable <- list(
		physical=goa.data,
		entityName="goa",
		entityDescription = "Bottom Trawl Data in the Gulf of Alaska",
		attributeList = attributeList
)


eml <- list(
           packageId = uuid::UUIDgenerate(),  
           system = "uuid",
           dataset = list(
               title = dataTitle,
               creator = wayne,
               pubDate = pubDate,
               # abstract = abstract,
               contact = wayne,
               additionalMetadata = list(metadata = list(unitList = unitList)),
               dataTable = goa.dataTable
               ))


# validate
eml_validate(eml)

# Write EML data file
write_eml(eml, file=here::here("metadata","goa.xml"))
```

The Ecological Metadata Language file can be accessed [here](https://github.com/pinskylab/OceanAdapt/blob/master/metaData/goa.xml)

