---
title: "Gulf of Mexico"
output: html_notebook
---
```{r setup, include = FALSE, message=FALSE}
# load libraries
library(EML)
library(emld)
library(tidyverse) # for data table wrangling
library(knitr) # for web presentation of tables
library(kableExtra) # for web presentation of tables

column_defs <- readRDS(here::here("metadata","column-definitions.rds"))
```
## Source:
[Gulf States Marine Fisheries Commission SEAMAP Groundfish Surveys](http://seamap.gsmfc.org)

## Related Papers: 
- [Gulf of Mexico Operations Manual](https://seamap.gsmfc.org/documents/SEAMAP%20Operations%20Manual%20March%202016.pdf)  
- [Derivation of Red Snapper Time Series from SEAMAP and Groundfish Trawl Surveys](http://sedarweb.org/docs/wpapers/SEDAR7_DW1.pdf)  
- [Comparisons of Relative Fishing Powers of Selected SEAMAP Survey Vessels](http://sedarweb.org/docs/wpapers/SEDAR7_DW53.pdf)  
- [Fishery-independent Bottom Trawl Surveys for Deep-water Fishes and Invertebrates of the U.S. Gulf of Mexico, 2002-08](https://spo.nmfs.noaa.gov/sites/default/files/pdf-content/MFR/mfr724/mfr7242.pdf)  
- [SEAMAP Oracle DMS definitions](http://seamap.gsmfc.org/documents/filedef.doc)  

## How we process the data:
- gmex_STAREC.csv has characters that prevent it from being parsed easily.  We attempt to make parsing more manageable by replacing the quotes in quoted character strings using the R regex below then reading the file in as a csv.
```{r eval=FALSE}
str_replace_all(gmex_station_raw, "\\\\\\\"", "\\\"\\\"")
```
- We only keep gear type "ST".  
- We trim out young of year records (only useful for count data) and those with UNKNOWN species.  
- We make two combined records where 'ANTHIAS TENUIS AND WOODSI', 'MOLLUSCA AND UNID.OTHER #01' share the same species code.  
- We trim to high quality SEAMAP summer trawls, based off the subset used by Jeff Rester's GS_TRAWL_05232011.sas.  
- We create a haulid by combining a 3 digit leading zero vessel number with a 3 digit leading zero cruise number and a 3 digit leading zero haul number, separated by "-", for example: (vessel-cruise-haul) 354-067-001.  
- We convert fathoms to meters. 
- We create a "strata" value by using lat, lon and depth to create a value in 100m bins.   
- We trim out or fix speed and duration records by trimming out tows of 0, >60, or unknown minutes.  
- We fix VESSEL_SP typo according to Jeff Rester: 30 = 3.  
- We trim out vessel speeds 0, unknown, or >5 (need vessel speed to calculate area trawled).  
- We remove a tow when paired tows exist, same lat/lon/year but different haulid.  
- We adjust wtcpue (biomass per standard tow) for area towed, in units of kg per 10,000 m2. Calculate area trawled in m2: knots * 1.8 km/hr/knot * 1000 m/km * minutes * 1 hr/60 min * width of gear in feet * 0.3 m/ft.
- We remove unidentified spp, white space only values, and adjust the following names:
```{r eval=FALSE}
    !spp %in% c('UNID CRUSTA', 'UNID OTHER', 'UNID.FISH', 'CRUSTACEA(INFRAORDER) BRACHYURA', 'MOLLUSCA AND UNID.OTHER #01', 'ALGAE', 'MISCELLANEOUS INVERTEBR', 'OTHER INVERTEBRATES')
  ) %>% 
  # adjust spp names
  mutate(
    spp = ifelse(GENUS_BGS == 'PELAGIA' & SPEC_BGS == 'NOCTUL', 'PELAGIA NOCTILUCA', spp), 
    BIO_BGS = ifelse(spp == "PELAGIA NOCTILUCA", 618030201, BIO_BGS), 
    spp = ifelse(GENUS_BGS == 'MURICAN' & SPEC_BGS == 'FULVEN', 'MURICANTHUS FULVESCENS', spp), 
    BIO_BGS = ifelse(spp == "MURICANTHUS FULVESCENS", 308011501, BIO_BGS), 
    spp = ifelse(grepl("APLYSIA", spp), "APLYSIA", spp), 
    spp = ifelse(grepl("AURELIA", spp), "AURELIA", spp), 
    spp = ifelse(grepl("BOTHUS", spp), "BOTHUS", spp), 
    spp = ifelse(grepl("CLYPEASTER", spp), "CLYPEASTER", spp), 
    spp = ifelse(grepl("CONUS", spp), "CONUS", spp), 
    spp = ifelse(grepl("CYNOSCION", spp), "CYNOSCION", spp), 
    spp = ifelse(grepl("ECHINASTER", spp), "ECHINASTER", spp),
    spp = ifelse(grepl("OPISTOGNATHUS", spp), "OPISTOGNATHUS", spp), 
    spp = ifelse(grepl("OPSANUS", spp), "OPSANUS", spp), 
    spp = ifelse(grepl("ROSSIA", spp), "ROSSIA", spp), 
    spp = ifelse(grepl("SOLENOCERA", spp), "SOLENOCERA", spp), 
    spp = ifelse(grepl("TRACHYPENEUS", spp), "TRACHYPENEUS", spp)
```
- We group the data by haulid, stratum, stratumarea, year, lat, lon, depth,  and spp and then sum up all of the wtcpue values for each group and reassign that as the wtcpue. 

## What the raw data includes:  
The current files of raw data for the Gulf of Mexico are `r list.files(here::here("data_raw"), pattern = "gmex")`.

### gmex_BGSREC.csv column definitions 
```{r include = FALSE}
sample_data <- read_csv(here::here("data_raw", "gmex_BGSREC.csv"), n_max = 6)
  
attributes <- tibble(
  attributeName = dput(names(sample_data))) %>% 
  left_join(column_defs) %>% 
  mutate(col_class = ifelse(is.na(col_class), "character", col_class), 
         attributeDefinition = ifelse(is.na(attributeDefinition), "unknown", attributeDefinition))

clas <- dput(attributes$col_class)

attributes <- attributes %>% 
  select(-col_class)

attributeList <- set_attributes(attributes, col_classes = clas)
```

```{r echo=FALSE, message=TRUE, warning=TRUE}
knitr::kable(attributes) %>% 
  kableExtra::kable_styling(latex_options = "scale_down")
```

### gmex_CRUISES.csv column definitions 
```{r include = FALSE}
sample_data <- read_csv("https://raw.githubusercontent.com/pinskylab/OceanAdapt/master/data_raw/gmex_CRUISES.csv", n_max = 6)
  
attributes <- tibble(
  attributeName = dput(names(sample_data))) %>% 
  left_join(column_defs) %>% 
  mutate(col_class = ifelse(is.na(col_class), "character", col_class), 
         attributeDefinition = ifelse(is.na(attributeDefinition), "unknown", attributeDefinition))

clas <- dput(attributes$col_class)

attributes <- attributes %>% 
  select(-col_class)

attributeList <- set_attributes(attributes, col_classes = clas)
```

```{r echo=FALSE, message=TRUE, warning=TRUE}
knitr::kable(attributes) %>% 
  kableExtra::kable_styling(latex_options = "scale_down")
```


### The remaining files are trawl data files, data added annually, with the column definitions
```{r include = FALSE}
# create a table with each column from the dataset represented as a row.  There will be an error message about undefined units but they are defined later when the dataset is created.
attributes <- tibble(
  attributeName = dput(names(sample_data))) %>% 
  left_join(column_defs) %>% 
  mutate(formatString = NA,
         formatString = ifelse(attributeName == "YEAR", unit, formatString), 
         formatString = ifelse(attributeName == "DATETIME", unit, formatString))

attributeList <- set_attributes(attributes, col_classes = c("numeric", "numeric", "character", "numeric", "Date", "Date", "numeric", "numeric", "character", "character", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric"))
```
```{r echo=FALSE}
knitr::kable(select(attributes, -formatString)) %>% 
  kableExtra::kable_styling(latex_options = "scale_down")
```

```{r include = FALSE}
# define custom units
id <- c("kilograms per tow", "number per tow", "square kilometers")
dimension <- c("weight", "count", "area")
unitTypes <- data.frame(
  id = id, dimension = dimension, stringsAsFactors = FALSE
)
units <- data.frame(
  id = id, unitType = unitTypes, stringsAsFactors = FALSE
)

unitList <- set_unitList(units, unitTypes)

# define the rest of the data
# <title>
	# character vector title
	dataTitle <- "Gulf of Mexico bottom trawl survey"

# <creator>
	# organization name
	afsc.name <- "National Oceanic and Atmospheric Administration (NOAA) Alaska Fisheries Science Center (AFSC) Resource Assessment and Conservation Engineering Division (RACE)"
	# create organization address
	afsc_address <- list( 
		deliveryPoint = "7600 Sand Point Way, N.E. bldg. 4",
		city = "Seattle",
		administrativeArea = "WA",
		postalCode = "98115",
		country = "USA"
	)
	
	publisher <- list(
	organizationName = afsc.name,
	address = afsc_address,
	onlineUrl = "http://www.afsc.noaa.gov/RACE/groundfish/survey_data/default.htm")

	R_person <- person(given = "Wayne", family = "Palsson", email = "Wayne.Palsson@noaa.gov", role = "cre")
	wayne <- as_emld(R_person)

	metadataProvider <- as_emld(person(given="Michelle", family = "Stuart",email = "michelle.stuart@rutgers.edu"))
	
others <- as.person("Bob Lauth [ctb] <Bob.Lauth@noaa.gov>")
associatedParty <- as_emld(others)
associatedParty[[1]]$organizationName <- afsc.name
	
# <pubDate>
pubDate <- "2012"

# <intellectualRights>
	# IR <- "this is just a long piece of text"

# <abstract>
abstract <- set_TextType(text = "The Resource Assessment and Conservation Engineering Division (RACE) of the Alaska Fisheries Science Center (AFSC) conducts bottom trawl surveys to monitor the condition of the demersal fish and crab stocks of Alaska. These data include catch per unit effort for each identified species at a standard set of stations. This is a subset of the main racebase datase. Excluded are non standard stations, earlier years using different gear, and other types of data collected other than species id, species weight, water temperature and depth.")

# This isn't working right now
# # <methods>
# 	methods <- 	list(dataSource="http://www.afsc.noaa.gov/RACE/groundfish/survey_data/default.htm",
# 				description="http://www.afsc.noaa.gov/RACE/groundfish/survey_data/metadata_template.php?fname=RACEweb.xml"
# 			) 
# 	set_methods(methods)

	

# ======================
# = Create EML Objects =
# ======================

goa.data <- set_physical("https://raw.githubusercontent.com/pinskylab/OceanAdapt/master/data_raw/ai1983_2000.csv")
# Create "dataTable": the physical data set with associated column definitions and units
goa.dataTable <- list(
		physical=goa.data,
		entityName="goa",
		entityDescription = "Bottom Trawl Data in the Gulf of Alaska",
		attributeList = attributeList
)


eml <- list(
           packageId = uuid::UUIDgenerate(),  
           system = "uuid",
           dataset = list(
               title = dataTitle,
               creator = wayne,
               pubDate = pubDate,
               # abstract = abstract,
               contact = wayne,
               additionalMetadata = list(metadata = list(unitList = unitList)),
               dataTable = goa.dataTable
               ))


# validate
eml_validate(eml)

# Write EML data file
write_eml(eml, file=here::here("metadata","goa.xml"))
```

The Ecological Metadata Language file can be accessed [here](https://github.com/pinskylab/OceanAdapt/blob/master/metaData/goa.xml)




