---
title: "Check species names"
output: html_notebook
---

Load libraries and data
```{r setup, include = FALSE}
#library(tidyverse)
library(dplyr)
library(plyr)
library(tidyr)
library(stringr)
library(readr)
library(rfishbase) # to get missing taxon data
library(janitor) # to clean names
library(ggplot2)
library(matrixStats)
library(gridExtra)
library(here) # to find yourself

# this is an R data of all of the individual regions
all_data <- load(here("data_clean", "individual-regions.rds"))

# this is the data after it has been joined with the spptaxonomy file but before trimming.
master_data <- readRDS(here("data_clean", "all-regions-full.rds"))

# this is the spptaxonomy file
tax <- read_csv(here("data_raw", "spptaxonomy.csv"))
```
```{r}

#create flag_spp function
flag_spp <- function(region){
    df <- region %>% 
      select(spp, year) %>% 
      group_by(spp) %>% 
      distinct()
    years <- unique(df$year)
    df2 <- data.frame(matrix(ncol = length(years)+1, nrow = length(df$spp)))
    x <- c('spp', unique(df$year))
    colnames(df2) <- x
    df2$spp <- df$spp
    
    n<-2
  for (year in years){
    df2[,n] <- c(apply(df, 1, function(r) any(r %in% c(year))))
    n <- n+1
    }
    df2 <- df2 %>% group_by(spp) %>% summarize_all(any)
    df3 <- data.frame(matrix(ncol = 3, nrow = length(df2$spp)))
    colnames(df3) <- c('spp', 'rlmax', 'PresToAbs')
    for (n in 1:length(df2$spp)){
      r <- rle(df2[n,])
      sp <- c(df2[n,1])
      rlmax <- max(r$lengths)
      df3[n,1] <- sp
      df3[n,2] <- rlmax
      df3[n,3] <- length(r$values)
    }
    x <- round(0.95*length(years))
    df3 <- df3[df3$rlmax < x,]
    df3 <- df3[df3$PresToAbs < 4,]
    df3
  
}

```
```{r}
#Flag spp for each region
flaglist <- list()

#Aleutian Islands
region <- ai
test <- flag_spp(region)

testdf <- data.frame(test$spp)
if (dim(testdf)[1] > 0) {
testdf$region <- 'Aleutian Islands'
} else {}
#write csv of flagged spp in region
write_csv(testdf, file = here::here("flagspp", "ai_flagspp.csv"))

  p1 <- region %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()

  test <- subset(region, spp %in% test$spp)
  
  p2 <- test %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()
  
  length(test$spp)/length(region$spp)
  #gmex, 1.04%; scot, 3.8%; ai, 4.99%; neusF, 2.2%; neusS, 2.84%; seusFALL, 1.9%; seusSPRING, 1.86%; seusSUMMER, 0.84%; wcann, 1.5%, wctri, 5.07%, ebs, 1.68%; goa, 1.61%
  temp <- grid.arrange(p1, p2, nrow = 2)
  ggsave(plot = temp, filename = here::here("plots", "ai_tax_check_test.png"))
  
#Canadian Pacific
region <- CPAC
test <- flag_spp(region)

testdf <- data.frame(test$spp)
if (dim(testdf)[1] > 0) {
testdf$region <- 'Canadian Pacific'
} else {}
#write csv of flagged spp in region
write_csv(testdf, file = here::here("flagspp", "cpac_flagspp.csv"))

  p1 <- region %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()

  test <- subset(region, spp %in% test$spp)
  
  p2 <- test %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()
  
  length(test$spp)/length(region$spp)
  #CPAC 3.6%
  temp <- grid.arrange(p1, p2, nrow = 2)
  ggsave(plot = temp, filename = here::here("plots", "cpac_tax_check_test.png"))
  

#Eastern Bering Sea
region <- ebs
test <- flag_spp(region)

testdf <- data.frame(test$spp)

if (dim(testdf)[1] > 0) {
testdf$region <- 'Eastern Bering Sea'
} else {}
#write csv of flagged spp in region
write_csv(testdf, file = here::here("flagspp", "ebs_flagspp.csv"))

  p1 <- region %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()

  test <- subset(region, spp %in% test$spp)
  
  p2 <- test %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()
  
  length(test$spp)/length(region$spp)
  #ebs, 1.74%
  temp <- grid.arrange(p1, p2, nrow = 2)
  ggsave(plot = temp, filename = here::here("plots", "ebs_tax_check_test.png"))

#Gulf of Mexico 
region <- gmex
test <- flag_spp(region)

testdf <- data.frame(test$spp)
if (dim(testdf)[1] > 0) {
testdf$region <- 'Gulf of Mexico'
} else {}
#write csv of flagged spp in region

  p1 <- region %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()

  #Remove species from the list that were first identified in 2008 (Jeff Rester said that they would likely have first appeared because of the expansion of the study area into the East Gulf, not because of any problem)
  test <- subset(region, spp %in% test$spp)
  test2 <- aggregate(year ~ spp, test, function(x) min(x))
  test3 <- test2[test2$year == 2008,]
  test4 <- test[!test$spp %in% test3$spp,]
  
  p2 <- test4 %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()
  
  length(test$spp)/length(region$spp)
  #gmex, 1.74%
  temp <- grid.arrange(p1, p2, nrow = 2)
  ggsave(plot = temp, filename = here::here("plots", "gmex_tax_check_test2.png"))
  
  testdf <- data.frame(unique(test4$spp))
if (dim(testdf)[1] > 0) {
testdf$region <- 'Gulf of Mexico'
} else {}
  write_csv(testdf, file = here::here("flagspp", "gmex_flagspp2.csv"))

#Gulf of Alaska
region <- goa
test <- flag_spp(region)

testdf <- data.frame(test$spp)
if (dim(testdf)[1] > 0) {
testdf$region <- 'Gulf of Alaska'
} else {}
#write csv of flagged spp in region
write_csv(testdf, file = here::here("flagspp", "goa_flagspp.csv"))

  p1 <- region %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()

  test <- subset(region, spp %in% test$spp)
  
  p2 <- test %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()
  
  length(test$spp)/length(region$spp)
  #goa, 1.62%
  temp <- grid.arrange(p1, p2, nrow = 2)
  ggsave(plot = temp, filename = here::here("plots", "goa_tax_check_test.png"))

#Gulf of St. Lawrence - North
region <- GSLnor
test <- flag_spp(region)

testdf <- data.frame(test$spp)
if (dim(testdf)[1] > 0) {
testdf$region <- 'Gulf of St. Lawrence North'
} else {}
#write csv of flagged spp in region
write_csv(testdf, file = here::here("flagspp", "GSLnor_flagspp.csv"))

  p1 <- region %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()

  test <- subset(region, spp %in% test$spp)
  
  p2 <- test %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()
  
  length(test$spp)/length(region$spp)
  #gmex, 1.04%; scot, 3.8%; ai, 4.99%; neusF, 2.2%; neusS, 2.84%; seusFALL, 1.9%; seusSPRING, 1.86%; seusSUMMER, 0.84%; wcann, 1.5%, wctri, 5.07%, ebs, 1.68%; goa, 1.61%
  temp <- grid.arrange(p1, p2, nrow = 2)
  ggsave(plot = temp, filename = here::here("plots", "GSLnor_tax_check_test.png"))

#Gulf of St. Lawrence - South
region <- GSLsouth
test <- flag_spp(region)

testdf <- data.frame(test$spp)
if (dim(testdf)[1] > 0) {
testdf$region <- 'Gulf of St. Lawrence South'
} else {}

#write csv of flagged spp in region
write_csv(testdf, file = here::here("flagspp", "GSLsouth_flagspp.csv"))

  p1 <- region %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()

  test <- subset(region, spp %in% test$spp)
  
  p2 <- test %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()
  
  length(test$spp)/length(region$spp)
  #gmex, 1.04%; scot, 3.8%; ai, 4.99%; neusF, 2.2%; neusS, 2.84%; seusFALL, 1.9%; seusSPRING, 1.86%; seusSUMMER, 0.84%; wcann, 1.5%, wctri, 5.07%, ebs, 1.68%; goa, 1.61%
  temp <- grid.arrange(p1, p2, nrow = 2)
  ggsave(plot = temp, filename = here::here("plots", "GSLsouth_tax_check_test.png"))  

  
#NEUS Fall
region <- neus_fall
test <- flag_spp(region)

testdf <- data.frame(test$spp)
if (dim(testdf)[1] > 0) {
testdf$region <- 'Northeast US Fall'
} else {}
#write csv of flagged spp in region
write_csv(testdf, file = here::here("flagspp", "neusF_flagspp.csv"))

  p1 <- region %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()

  test <- subset(region, spp %in% test$spp)
  
  p2 <- test %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()
  
  length(test$spp)/length(region$spp)
  #gmex, 1.04%; scot, 3.8%; ai, 4.99%; neusF, 2.2%; neusS, 2.84%; seusFALL, 1.9%; seusSPRING, 1.86%; seusSUMMER, 0.84%; wcann, 1.5%, wctri, 5.07%, ebs, 1.68%; goa, 1.61%
  temp <- grid.arrange(p1, p2, nrow = 2)
  ggsave(plot = temp, filename = here::here("plots", "neusF_tax_check_test.png"))

#NEUS Spring  
region <- neus_spring
test <- flag_spp(region)

testdf <- data.frame(test$spp)
if (dim(testdf)[1] > 0) {
testdf$region <- 'Northeast US Spring'
} else {}
#write csv of flagged spp in region
write_csv(testdf, file = here::here("flagspp", "neusS_flagspp.csv"))

  p1 <- region %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()

  test <- subset(region, spp %in% test$spp)
  
  p2 <- test %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()
  
  length(test$spp)/length(region$spp)
  #gmex, 1.04%; scot, 3.8%; ai, 4.99%; neusF, 2.2%; neusS, 2.84%; seusFALL, 1.9%; seusSPRING, 1.86%; seusSUMMER, 0.84%; wcann, 1.5%, wctri, 5.07%, ebs, 1.68%; goa, 1.61%
  temp <- grid.arrange(p1, p2, nrow = 2)
  ggsave(plot = temp, filename = here::here("plots", "neusS_tax_check_test.png"))
  
# Maritimes
region <- mar
test <- flag_spp(region)

testdf <- data.frame(test$spp)
if (dim(testdf)[1] > 0) {
testdf$region <- 'Maritimes'
} else {}
#write csv of flagged spp in region
write_csv(testdf, file = here::here("flagspp", "mar_flagspp.csv"))

  p1 <- region %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()

  test <- subset(region, spp %in% test$spp)
  
  p2 <- test %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()
  
  length(test$spp)/length(region$spp)
  #gmex, 1.04%; scot, 3.8%; ai, 4.99%; neusF, 2.2%; neusS, 2.84%; seusFALL, 1.9%; seusSPRING, 1.86%; seusSUMMER, 0.84%; wcann, 1.5%, wctri, 5.07%, ebs, 1.68%; goa, 1.61%
  temp <- grid.arrange(p1, p2, nrow = 2)
  ggsave(plot = temp, filename = here::here("plots", "mar_tax_check_test.png"))

#SEUS Fall
region <- seusFALL
test <- flag_spp(region)

testdf <- data.frame(test$spp)
if (dim(testdf)[1] > 0) {
testdf$region <- 'Southeast US Fall'
} else {}
#write csv of flagged spp in region
write_csv(testdf, file = here::here("flagspp", "seusFALL_flagspp.csv"))

  p1 <- region %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()

  test <- subset(region, spp %in% test$spp)
  
  p2 <- test %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()
  
  length(test$spp)/length(region$spp)
  #gmex, 1.04%; scot, 3.8%; ai, 4.99%; neusF, 2.2%; neusS, 2.84%; seusFALL, 1.9%; seusSPRING, 1.86%; seusSUMMER, 0.84%; wcann, 1.5%, wctri, 5.07%, ebs, 1.68%; goa, 1.61%
  temp <- grid.arrange(p1, p2, nrow = 2)
  ggsave(plot = temp, filename = here::here("plots", "seusFALL_tax_check_test.png"))

#SEUS Spring
region <- seusSPRING
test <- flag_spp(region)

testdf <- data.frame(test$spp)
if (dim(testdf)[1] > 0) {
testdf$region <- 'Southeast US Spring'
} else {}
#write csv of flagged spp in region
write_csv(testdf, file = here::here("flagspp", "seusSPRING_flagspp.csv"))

  p1 <- region %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()

  test <- subset(region, spp %in% test$spp)
  
  p2 <- test %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()
  
  length(test$spp)/length(region$spp)
  #gmex, 1.04%; scot, 3.8%; ai, 4.99%; neusF, 2.2%; neusS, 2.84%; seusFALL, 1.9%; seusSPRING, 1.86%; seusSUMMER, 0.84%; wcann, 1.5%, wctri, 5.07%, ebs, 1.68%; goa, 1.61%
  temp <- grid.arrange(p1, p2, nrow = 2)
  ggsave(plot = temp, filename = here::here("plots", "seusSPRING_tax_check_test.png"))

#SEUS Summer
region <- seusSUMMER
test <- flag_spp(region)

testdf <- data.frame(test$spp)
if (dim(testdf)[1] > 0) {
testdf$region <- 'Southeast US Summer'
} else {}
#write csv of flagged spp in region
write_csv(testdf, file = here::here("flagspp", "seusSUMMER_flagspp.csv"))

  p1 <- region %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()

  test <- subset(region, spp %in% test$spp)
  
  p2 <- test %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()
  
  length(test$spp)/length(region$spp)
  #gmex, 1.04%; scot, 3.8%; ai, 4.99%; neusF, 2.2%; neusS, 2.84%; seusFALL, 1.9%; seusSPRING, 1.86%; seusSUMMER, 0.84%; wcann, 1.5%, wctri, 5.07%, ebs, 1.68%; goa, 1.61%
  temp <- grid.arrange(p1, p2, nrow = 2)
  ggsave(plot = temp, filename = here::here("plots", "seusSUMMER_tax_check_test.png"))
  
#West Coast Annual
region <- wcann
test <- flag_spp(region)

testdf <- data.frame(test$spp)
if (dim(testdf)[1] > 0) {
testdf$region <- 'Westcoast Annual'
} else {}
#write csv of flagged spp in region
write_csv(testdf, file = here::here("flagspp", "wcann_flagspp.csv"))

  p1 <- region %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()

  test <- subset(region, spp %in% test$spp)
  
  p2 <- test %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()
  
  length(test$spp)/length(region$spp)
  #gmex, 1.04%; scot, 3.8%; ai, 4.99%; neusF, 2.2%; neusS, 2.84%; seusFALL, 1.9%; seusSPRING, 1.86%; seusSUMMER, 0.84%; wcann, 1.5%, wctri, 5.07%, ebs, 1.68%; goa, 1.61%
  temp <- grid.arrange(p1, p2, nrow = 2)
  ggsave(plot = temp, filename = here::here("plots", "wcann_tax_check_test.png"))
  
#AWest Coast Triannual
region <- wctri
test <- flag_spp(region)

testdf <- data.frame(test$spp)
if (dim(testdf)[1] > 0) {
testdf$region <- 'Westcoast Triennial'
} else {}
#write csv of flagged spp in region
write_csv(testdf, file = here::here("flagspp", "wctri_flagspp.csv"))

  p1 <- region %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()

  test <- subset(region, spp %in% test$spp)
  
  p2 <- test %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()
  
  length(test$spp)/length(region$spp)
  #gmex, 1.04%; scot, 3.8%; ai, 4.99%; neusF, 2.2%; neusS, 2.84%; seusFALL, 1.9%; seusSPRING, 1.86%; seusSUMMER, 0.84%; wcann, 1.5%, wctri, 5.07%, ebs, 1.68%; goa, 1.61%
  temp <- grid.arrange(p1, p2, nrow = 2)
  ggsave(plot = temp, filename = here::here("plots", "wctri_tax_check_test.png"))

```


Aurore asked about how the taxonomic file does not cover all 'spp' names from the surveys and ended up with species names without their corresponding 'taxon' in the file 'spptaxonomy.csv' (>800), and include several fish species (such as winter flounder and several sebastes species) from several surveys (EBS, GOA, NEUS, WCTRI, WCANN, GMEX).

# examine column names in all-regions-full.RData
```{r}
names(all_data)
```
# Combine the imported and cleaned regions and check the values before they are trimmed to the spptaxonomy file
```{r}
#load( here("data_clean", "individual-regions.rds"))
#all_regions <- rbind(ai, CPAC, ebs, gmex, goa,GSLnor, GSLsouth, mar, neus_fall, neus_spring, seusFALL, seusSPRING, seusSUMMER, wcann, wctri)

#names(all_regions)



all_regions <- readRDS( here("data_clean", "all-regions-trimmed.rds"))

names(all_regions)

```

# examine only the spp column, only unique values, and assign a test id number
```{r}

all_regions_spp <- all_regions %>% 
  select(spp) %>% 
  distinct() %>% 
  mutate(spp_id = 1:nrow(.))

```
# Anti-join this spp list to the taxon column from the spptaxonomy file to see which spp are not represented there
```{r}

not_in_tax <- anti_join(all_regions_spp, tax, by = c("spp"="taxon")) %>% 
  # format the column into genus and spp so that I can check it against fishbase
  separate(spp, into= c("genus", "spp", "extra", "extra1"), sep = " ") %>%
  mutate(genus = str_to_title(genus)) %>% 
  select(-contains("extra")) 
```
Check that all whitespace was removed from names
```{r, eval = FALSE}
test <- not_in_tax %>% 
  separate(spp, into = c("spp", "white"), sep = " ") %>% 
  filter(!is.na(white)) 

# should return one row for the spp with 2 names
```

# Check against fishbase
```{r}
not_in_tax <- not_in_tax %>% 
  mutate(name = str_c(genus, spp, sep = " "))

fb_check <- fishbase %>% 
  # join the genus and spp columns into one
  mutate(name = str_c(Genus, Species, sep = " ")) %>% 
  filter(name %in% not_in_tax$name) %>% 
  clean_names() %>% 
  select(-contains("code"), -species_ref_no) %>% 
  dplyr::rename(common = f_bname) 
#155 in fb check  

sb_check <- sealifebase %>% 
  mutate(name = str_c(Genus, Species, sep = " ")) %>% 
  filter(name %in% not_in_tax$name) %>% 
  clean_names() %>% 
  select(-contains("code"), -species_ref_no) %>% 
  dplyr::rename(common = f_bname) 
# 54 in sb_check
```

# Add the results of species to the spptaxonomy.csv
```{r}
col_spec <- cols(.default = col_character())
tax_raw <- read_csv(here("data_raw", "spptaxonomy.csv"), col_types = col_spec)

# add columns to fb_check to match tax_raw
names(tax_raw)

# have to add the taxon back into fb_check based on the species id I created earlier
matching <- left_join(not_in_tax, select(fb_check, -genus, -species), by = c("name")) %>% 
  filter(!is.na(family)) %>% 
  dplyr::rename(species = spp) %>% 
  left_join(all_regions_spp, by = "spp_id") %>% 
  dplyr::rename(taxon = spp) %>% 
  select(-spp_id, -sub_family) %>% 
  mutate(superclass = NA, 
         subphylum = NA, 
         phylum = "Chordata", 
         kingdom = "Animalia")

names(matching)
names(tax_raw)

tax_added <- rbind(tax_raw, matching)
```

#add sealifebase names
```{r}
names(tax_added)
names(sb_check)

matching <- left_join(not_in_tax, select(sb_check, -genus, -species), by = c("name")) %>% 
  filter(!is.na(family)) %>% 
  dplyr::rename(species = spp) %>% 
  left_join(all_regions_spp, by = "spp_id") %>% 
  dplyr::rename(taxon = spp) %>% 
  select(-spp_id, -sub_family) %>% 
  mutate(superclass = NA, 
         subphylum = NA, 
         phylum = "Chordata", 
         kingdom = "Animalia")

tax_added <- rbind(tax_added, matching)
```

# There are still genera only rows that can be matched
```{r}

gen_not_in_tax <- anti_join(not_in_tax, select(matching, -genus), by = "name")

gen_fb_check <- fishbase %>% 
  # join the genus and spp columns into one
  filter(Genus %in% gen_not_in_tax$genus) %>% 
  clean_names() %>% 
  select(-contains("code"), -species_ref_no, -species, -f_bname) %>% 
  distinct()
  
# make sure the same genus name is not matching more than one family
test <- gen_fb_check %>% 
  group_by(genus) %>% 
  filter(n() > 1)


gen_sb_check <- sealifebase %>% 
  filter(Genus %in% gen_not_in_tax$genus) %>% 
  clean_names() %>% 
  select(-contains("code"), -species_ref_no, -species, -f_bname) %>% 
  distinct()

# make sure the same genus name is not matching more than one family
test <- gen_sb_check %>% 
  group_by(genus) %>% 
  filter(n() > 1)

gen_check <- rbind(gen_fb_check, gen_sb_check)
```

# Add the results of genera to the spptaxonomy.csv
```{r}
# add columns to fb_check to match tax_raw
names(tax_added)

# have to add the taxon back into fb_check based on the species id I created earlier
matching <- left_join(not_in_tax, gen_check, by = "genus") %>% 
  filter(!is.na(family), is.na(spp)) %>% 
  dplyr::rename(species = spp) %>% 
  left_join(all_regions_spp, by = "spp_id") %>% 
  dplyr::rename(taxon = spp) %>% 
  select(-spp_id, -sub_family) %>% 
  mutate(superclass = NA, 
         subphylum = NA, 
         phylum = NA, 
         kingdom = NA, 
         common = NA)

names(tax_added)
names(matching)

tax_added <- rbind(tax_added, matching)

# anything left?
not_in_tax <- anti_join(all_regions_spp, tax_added, by = c("spp"="taxon"))
#103 taxon names left, most corresponse to families, some unusal designations
```


# Write to csv
```{r}
write_csv(tax_added, here("data_raw", "spptaxonomy.csv"))
```

# Write to csv
```{r}
library(compare)
comparison <- compare(tax_raw,tax_added,allowAll=TRUE)
```

Malin asked if there were other Echinoidea names that don’t have “crushed urchin"

```{r eval=FALSE}
echinoidea <- all_data %>% 
  filter(grepl("Echinoid", spp)) %>% 
  select(spp, common) %>% 
  distinct()

other <- all_data %>% 
  filter(grepl("spp", spp)) %>% 
  select(spp, common) %>% 
  distinct()

```

Daniel asked if there is a fish called the Alantic seabob in GOM - yep, there is a type-o, fixing it here
```{r eval=FALSE}
# this is the spptaxonomy file
tax <- read_csv(here("data_raw", "spptaxonomy.csv"))

master_data %>% 
  filter(grepl("Alantic", common))

new <- tax %>% 
  mutate(common = ifelse(grepl("Alantic", common), "Atlantic seabob", common))

tax %>% 
  filter(grepl("Alantic", common))

new %>% 
  filter(grepl("Alantic", common))


write_csv(new, here("data_raw", "spptaxonomy.csv"))
```

