---
title: "Check species names"
output: html_notebook
---

Load libraries and data
```{r setup, include = FALSE}
#library(tidyverse)
library(dplyr)
library(plyr)
library(tidyr)
library(stringr)
library(readr)
library(rfishbase) # to get missing taxon data
library(janitor) # to clean names
library(ggplot2)
library(matrixStats)
library(gridExtra)
library(here) # to find yourself

# this is an R data of all of the individual regions
all_data <- load(here("data_clean", "individual-regions.rda"))

# this is the data after it has been joined with the spptaxonomy file but before trimming.
master_data <- readRDS(here("data_clean", "all-regions-full.rds"))

# this is the spptaxonomy file
tax <- read_csv(here("data_raw", "spptaxonomy.csv"))
```
```{r}
flag_spp <- function(region){
    df <- region %>% 
      select(spp, year) %>% 
      group_by(spp) %>% 
      distinct()
    years <- unique(df$year)
    df2 <- data.frame(matrix(ncol = length(years)+1, nrow = length(df$spp)))
    x <- c('spp', unique(df$year))
    colnames(df2) <- x
    df2$spp <- df$spp
    
    n<-2
  for (year in years){
    df2[,n] <- c(apply(df, 1, function(r) any(r %in% c(year))))
    n <- n+1
    }
    df2 <- df2 %>% group_by(spp) %>% summarize_all(any)
    df3 <- data.frame(matrix(ncol = 3, nrow = length(df2$spp)))
    colnames(df3) <- c('spp', 'rlmax', 'PresToAbs')
    for (n in 1:length(df2$spp)){
      r <- rle(df2[n,])
      sp <- c(df2[n,1])
      rlmax <- max(r$lengths)
      df3[n,1] <- sp
      df3[n,2] <- rlmax
      df3[n,3] <- length(r$values)
    }
    x <- round(0.95*length(years))
    df3 <- df3[df3$rlmax < x,]
    df3 <- df3[df3$PresToAbs < 4,]
    df3
  
}
```
```{r}
#Flag spp for each region
flaglist <- list()

#Aleutian Islands
region <- ai
test <- flag_spp(region)

testdf <- data.frame(test$spp)
#write csv of flagged spp in region
write_csv(testdf, file = here::here("flagspp", "ai_flagspp.csv"))

  p1 <- region %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()

  test <- subset(region, spp %in% test$spp)
  
  p2 <- test %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()
  
  length(test$spp)/length(region$spp)
  #gmex, 1.04%; scot, 3.8%; ai, 4.99%; neusF, 2.2%; neusS, 2.84%; seusFALL, 1.9%; seusSPRING, 1.86%; seusSUMMER, 0.84%; wcann, 1.5%, wctri, 5.07%, ebs, 1.68%; goa, 1.61%
  temp <- grid.arrange(p1, p2, nrow = 2)
  ggsave(plot = temp, filename = here::here("plots", "ai_tax_check_test.png"))
  
test

#Eastern Bering Sea
region <- ebs
test <- flag_spp(region)

testdf <- data.frame(test$spp)
#write csv of flagged spp in region
write_csv(testdf, file = here::here("flagspp", "ebs_flagspp.csv"))

  p1 <- region %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()

  test <- subset(region, spp %in% test$spp)
  
  p2 <- test %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()
  
  length(test$spp)/length(region$spp)
  #gmex, 1.04%; scot, 3.8%; ai, 4.99%; neusF, 2.2%; neusS, 2.84%; seusFALL, 1.9%; seusSPRING, 1.86%; seusSUMMER, 0.84%; wcann, 1.5%, wctri, 5.07%, ebs, 1.68%; goa, 1.61%
  temp <- grid.arrange(p1, p2, nrow = 2)
  ggsave(plot = temp, filename = here::here("plots", "ebs_tax_check_test.png"))

#Gulf of Mexico 
  region <- gmex
test <- flag_spp(region)


  p1 <- region %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()

  test <- subset(region, spp %in% test$spp)
  
  p2 <- test %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()
  
  length(test$spp)/length(region$spp)
  #gmex, 1.04%; scot, 3.8%; ai, 4.99%; neusF, 2.2%; neusS, 2.84%; seusFALL, 1.9%; seusSPRING, 1.86%; seusSUMMER, 0.84%; wcann, 1.5%, wctri, 5.07%, ebs, 1.68%; goa, 1.61%
  temp <- grid.arrange(p1, p2, nrow = 2)
  ggsave(plot = temp, filename = here::here("plots", "gmex_tax_check_test.png"))

#Gulf of Alaska
region <- goa
test <- flag_spp(region)

testdf <- data.frame(test$spp)
#write csv of flagged spp in region
write_csv(testdf, file = here::here("flagspp", "goa_flagspp.csv"))

  p1 <- region %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()

  test <- subset(region, spp %in% test$spp)
  
  p2 <- test %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()
  
  length(test$spp)/length(region$spp)
  #gmex, 1.04%; scot, 3.8%; ai, 4.99%; neusF, 2.2%; neusS, 2.84%; seusFALL, 1.9%; seusSPRING, 1.86%; seusSUMMER, 0.84%; wcann, 1.5%, wctri, 5.07%, ebs, 1.68%; goa, 1.61%
  temp <- grid.arrange(p1, p2, nrow = 2)
  ggsave(plot = temp, filename = here::here("plots", "goa_tax_check_test.png"))

#NEUS Fall
region <- neusF
test <- flag_spp(region)

testdf <- data.frame(test$spp)
#write csv of flagged spp in region
write_csv(testdf, file = here::here("flagspp", "neusF_flagspp.csv"))

  p1 <- region %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()

  test <- subset(region, spp %in% test$spp)
  
  p2 <- test %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()
  
  length(test$spp)/length(region$spp)
  #gmex, 1.04%; scot, 3.8%; ai, 4.99%; neusF, 2.2%; neusS, 2.84%; seusFALL, 1.9%; seusSPRING, 1.86%; seusSUMMER, 0.84%; wcann, 1.5%, wctri, 5.07%, ebs, 1.68%; goa, 1.61%
  temp <- grid.arrange(p1, p2, nrow = 2)
  ggsave(plot = temp, filename = here::here("plots", "neusF_tax_check_test.png"))

#NEUS Summer  
region <- neusS
test <- flag_spp(region)

testdf <- data.frame(test$spp)
#write csv of flagged spp in region
write_csv(testdf, file = here::here("flagspp", "neusS_flagspp.csv"))

  p1 <- region %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()

  test <- subset(region, spp %in% test$spp)
  
  p2 <- test %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()
  
  length(test$spp)/length(region$spp)
  #gmex, 1.04%; scot, 3.8%; ai, 4.99%; neusF, 2.2%; neusS, 2.84%; seusFALL, 1.9%; seusSPRING, 1.86%; seusSUMMER, 0.84%; wcann, 1.5%, wctri, 5.07%, ebs, 1.68%; goa, 1.61%
  temp <- grid.arrange(p1, p2, nrow = 2)
  ggsave(plot = temp, filename = here::here("plots", "neusS_tax_check_test.png"))
  
#Scotian Shelf (soon to be replaced by Maritimes)
region <- scot
test <- flag_spp(region)

testdf <- data.frame(test$spp)
#write csv of flagged spp in region
write_csv(testdf, file = here::here("flagspp", "scot_flagspp.csv"))

  p1 <- region %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()

  test <- subset(region, spp %in% test$spp)
  
  p2 <- test %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()
  
  length(test$spp)/length(region$spp)
  #gmex, 1.04%; scot, 3.8%; ai, 4.99%; neusF, 2.2%; neusS, 2.84%; seusFALL, 1.9%; seusSPRING, 1.86%; seusSUMMER, 0.84%; wcann, 1.5%, wctri, 5.07%, ebs, 1.68%; goa, 1.61%
  temp <- grid.arrange(p1, p2, nrow = 2)
  ggsave(plot = temp, filename = here::here("plots", "scot_tax_check_test.png"))

#SEUS Fall
region <- seusFALL
test <- flag_spp(region)

testdf <- data.frame(test$spp)
#write csv of flagged spp in region
write_csv(testdf, file = here::here("flagspp", "seusFALL_flagspp.csv"))

  p1 <- region %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()

  test <- subset(region, spp %in% test$spp)
  
  p2 <- test %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()
  
  length(test$spp)/length(region$spp)
  #gmex, 1.04%; scot, 3.8%; ai, 4.99%; neusF, 2.2%; neusS, 2.84%; seusFALL, 1.9%; seusSPRING, 1.86%; seusSUMMER, 0.84%; wcann, 1.5%, wctri, 5.07%, ebs, 1.68%; goa, 1.61%
  temp <- grid.arrange(p1, p2, nrow = 2)
  ggsave(plot = temp, filename = here::here("plots", "seusFALL_tax_check_test.png"))

#SEUS Spring
region <- seusSPRING
test <- flag_spp(region)

testdf <- data.frame(test$spp)
#write csv of flagged spp in region
write_csv(testdf, file = here::here("flagspp", "seusSPRING_flagspp.csv"))

  p1 <- region %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()

  test <- subset(region, spp %in% test$spp)
  
  p2 <- test %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()
  
  length(test$spp)/length(region$spp)
  #gmex, 1.04%; scot, 3.8%; ai, 4.99%; neusF, 2.2%; neusS, 2.84%; seusFALL, 1.9%; seusSPRING, 1.86%; seusSUMMER, 0.84%; wcann, 1.5%, wctri, 5.07%, ebs, 1.68%; goa, 1.61%
  temp <- grid.arrange(p1, p2, nrow = 2)
  ggsave(plot = temp, filename = here::here("plots", "seusSPRING_tax_check_test.png"))

#SEUS Summer
region <- seusSUMMER
test <- flag_spp(region)

testdf <- data.frame(test$spp)
#write csv of flagged spp in region
write_csv(testdf, file = here::here("flagspp", "seusSUMMER_flagspp.csv"))

  p1 <- region %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()

  test <- subset(region, spp %in% test$spp)
  
  p2 <- test %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()
  
  length(test$spp)/length(region$spp)
  #gmex, 1.04%; scot, 3.8%; ai, 4.99%; neusF, 2.2%; neusS, 2.84%; seusFALL, 1.9%; seusSPRING, 1.86%; seusSUMMER, 0.84%; wcann, 1.5%, wctri, 5.07%, ebs, 1.68%; goa, 1.61%
  temp <- grid.arrange(p1, p2, nrow = 2)
  ggsave(plot = temp, filename = here::here("plots", "seusSUMMER_tax_check_test.png"))
  
#West Coast Annual
region <- wcann
test <- flag_spp(region)

testdf <- data.frame(test$spp)
#write csv of flagged spp in region
write_csv(testdf, file = here::here("flagspp", "wcann_flagspp.csv"))

  p1 <- region %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()

  test <- subset(region, spp %in% test$spp)
  
  p2 <- test %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()
  
  length(test$spp)/length(region$spp)
  #gmex, 1.04%; scot, 3.8%; ai, 4.99%; neusF, 2.2%; neusS, 2.84%; seusFALL, 1.9%; seusSPRING, 1.86%; seusSUMMER, 0.84%; wcann, 1.5%, wctri, 5.07%, ebs, 1.68%; goa, 1.61%
  temp <- grid.arrange(p1, p2, nrow = 2)
  ggsave(plot = temp, filename = here::here("plots", "wcann_tax_check_test.png"))
  
#AWest Coast Triannual
region <- wctri
test <- flag_spp(region)

testdf <- data.frame(test$spp)
#write csv of flagged spp in region
write_csv(testdf, file = here::here("flagspp", "wctri_flagspp.csv"))

  p1 <- region %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()

  test <- subset(region, spp %in% test$spp)
  
  p2 <- test %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_point()
  
  length(test$spp)/length(region$spp)
  #gmex, 1.04%; scot, 3.8%; ai, 4.99%; neusF, 2.2%; neusS, 2.84%; seusFALL, 1.9%; seusSPRING, 1.86%; seusSUMMER, 0.84%; wcann, 1.5%, wctri, 5.07%, ebs, 1.68%; goa, 1.61%
  temp <- grid.arrange(p1, p2, nrow = 2)
  ggsave(plot = temp, filename = here::here("plots", "wctri_tax_check_test.png"))

```


Aurore asked about how the taxonomic file does not cover all 'spp' names from the surveys and ended up with species names without their corresponding 'taxon' in the file 'spptaxonomy.csv' (>800), and include several fish species (such as winter flounder and several sebastes species) from several surveys (EBS, GOA, NEUS, WCTRI, WCANN, GMEX).

# examine column names in all-regions-full.RData
```{r}
names(all_data)
```
# Combine the imported and cleaned regions and check the values before they are trimmed to the spptaxonomy file
```{r}
load( here("data_clean", "individual-regions.rda"))
all_regions <- rbind(ai, ebs, gmex, goa, neusF, neusS, scot, seusFALL, seusSPRING, seusSUMMER, wcann, wctri)

names(all_regions)

```

# examine only the spp column, only unique values, and assign a test id number
```{r}

all_regions_spp <- all_regions %>% 
  select(spp) %>% 
  distinct() %>% 
  mutate(spp_id = 1:nrow(.))
  
```
# Anti-join this spp list to the taxon column from the spptaxonomy file to see which spp are not represented there
```{r}
not_in_tax <- anti_join(all_regions_spp, tax, by = c("spp"="taxon")) %>% 
  # format the column into genus and spp so that I can check it against fishbase
  separate(spp, into= c("genus", "spp", "extra", "extra1"), sep = " ") %>%
  mutate(genus = str_to_title(genus), 
         spp = str_c(spp, extra, sep = " "),
         spp = str_to_lower(spp), 
         spp = str_trim(spp)) %>% 
  select(-contains("extra")) 


```
Check that all whitespace was removed from names
```{r, eval = FALSE}
test <- not_in_tax %>% 
  separate(spp, into = c("spp", "white"), sep = " ") %>% 
  filter(!is.na(white)) 
# should return one row for the spp with 2 names
```

# Check against fishbase
```{r}
not_in_tax <- not_in_tax %>% 
  mutate(name = str_c(genus, spp, sep = " "))

fb_check <- fishbase %>% 
  # join the genus and spp columns into one
  mutate(name = str_c(Genus, Species, sep = " ")) %>% 
  filter(name %in% not_in_tax$name) %>% 
  clean_names() %>% 
  select(-contains("code"), -species_ref_no) %>% 
  rename(common = f_bname) 
  

sb_check <- sealifebase %>% 
  mutate(name = str_c(Genus, Species, sep = " ")) %>% 
  filter(name %in% not_in_tax$name)
# none in sealife base
```

# Add the results of species to the spptaxonomy.csv
```{r}
col_spec <- cols(.default = col_character())
tax_raw <- read_csv(here("data_raw", "spptaxonomy.csv"), col_types = col_spec)

# add columns to fb_check to match tax_raw
names(tax_raw)

# have to add the taxon back into fb_check based on the species id I created earlier
matching <- left_join(not_in_tax, select(fb_check, -genus, -species), by = c("name")) %>% 
  filter(!is.na(family)) %>% 
  rename(species = spp) %>% 
  left_join(all_regions_spp, by = "spp_id") %>% 
  rename(taxon = spp) %>% 
  select(-spp_id, -sub_family) %>% 
  mutate(superclass = NA, 
         subphylum = NA, 
         phylum = "Chordata", 
         kingdom = "Animalia")

names(matching)

tax_added <- rbind(tax_raw, matching)

```
# There are still genera only rows that can be matched
```{r}
gen_not_in_tax <- anti_join(not_in_tax, select(matching, -genus),by = "name") %>% 
  filter(is.na(spp))

gen_fb_check <- fishbase %>% 
  # join the genus and spp columns into one
  filter(Genus %in% gen_not_in_tax$genus) %>% 
  clean_names() %>% 
  select(-contains("code"), -species_ref_no, -species, -f_bname) %>% 
  distinct()
  
# make sure the same genus name is not matching more than one family
test <- gen_fb_check %>% 
  group_by(genus) %>% 
  filter(n() > 1)


gen_sb_check <- sealifebase %>% 
  filter(Genus %in% gen_not_in_tax$genus) %>% 
  clean_names() %>% 
  select(-contains("code"), -species_ref_no, -species, -f_bname) %>% 
  distinct()

# make sure the same genus name is not matching more than one family
test <- gen_sb_check %>% 
  group_by(genus) %>% 
  filter(n() > 1)

gen_check <- rbind(gen_fb_check, gen_sb_check)
```

# Add the results of genera to the spptaxonomy.csv
```{r}
# add columns to fb_check to match tax_raw
names(tax_added)

# have to add the taxon back into fb_check based on the species id I created earlier
matching <- left_join(not_in_tax, gen_check, by = "genus") %>% 
  filter(!is.na(family), is.na(spp)) %>% 
  rename(species = spp) %>% 
  left_join(all_regions_spp, by = "spp_id") %>% 
  rename(taxon = spp) %>% 
  select(-spp_id, -sub_family) %>% 
  mutate(superclass = NA, 
         subphylum = NA, 
         phylum = NA, 
         kingdom = NA, 
         common = NA)

names(tax_added)
names(matching)

tax_added <- rbind(tax_added, matching)

# anything left?
not_in_tax <- anti_join(all_regions_spp, tax_added, by = c("spp"="taxon"))

```
```{r}
 p1 <- wcann %>%
    select(spp, year) %>%
    ggplot(aes(x = year, y = spp)) +
    geom_jitter()

  test <- wcann %>% 
    select(spp, year) %>% 
    group_by(spp) #%>% 
    #summarise(count = n())
  
  unique(test$year)

```

# Write to csv
```{r}
write_csv(tax_added, here("data_raw", "spptaxonomy.csv"))
```

# Write to csv
```{r}
library(compare)
comparison <- compare(tax_raw,tax_added,allowAll=TRUE)
```

Malin asked if there were other Echinoidea names that don’t have “crushed urchin"

```{r eval=FALSE}
echinoidea <- all_data %>% 
  filter(grepl("Echinoid", spp)) %>% 
  select(spp, common) %>% 
  distinct()

other <- all_data %>% 
  filter(grepl("spp", spp)) %>% 
  select(spp, common) %>% 
  distinct()

```

Daniel asked if there is a fish called the Alantic seabob in GOM - yep, there is a type-o, fixing it here
```{r eval=FALSE}
master_data %>% 
  filter(grepl("Alantic", common))

new <- tax %>% 
  mutate(common = ifelse(grepl("Alantic", common), "Atlantic seabob", common))

tax %>% 
  filter(grepl("Alantic", common))

new %>% 
  filter(grepl("Alantic", common))


write_csv(new, here("data_raw", "spptaxonomy.csv"))
```

